\bf Service Chain Placement in SDNs \thanks{ A preliminary version appeared in the 3rd International Workshop on Algorithmic Aspects of Cloud Computing (ALGOCLOUD), 2017. Research supported in part by Network Programming (Neptune) Consortium, Israel.}

### 2019-01-28 src/chain.tex
fix

% DEBUG
%\usepackage{lipsum}
%\usepackage{todonotes}
%\usepackage{lineno}
%\linenumbers

@@

In this paper we consider an SDN network that employs
NVF~\cite{SDN-NFV15}.  Such networks attracted a lot of attention in
the networking community, especially from a systems point of view
(see, e.g.,~\cite{GVPGKDA14,HVSBFTF15}).  Several papers also
considered the algorithmic aspects of such networks but mainly
presented heuristics (see, e.g,~\cite{soule2014merlin}).  For more
details see~\cite{ERS16} and references therein.

@@

\subsection{Model}
%
An instance of the \scplong (\scp) problem is composed of three
components, a physical network, a virtual specification, and placement
costs:
\begin{description}
\item[Physical network:]
%
  The physical network is a graph $G = (V,E)$.  Each node $v \in V$
  has a non-negative processing capacity $p(v)$, and each directed
  edge $e \in E$ has a non-negative bandwidth capacity $b(e)$.
%
  Without loss of generality, we assume that $p(s) = p(t) = 0$.

@@

\item[Virtual specification:]
%
  The description of a request for a service chain consists of a
  physical source $s \in V$, a physical destination $t \in V$, and a
  directed acyclic graph (DAG) $\calG = (\calV,\calE)$.
%
  Without loss of generality, we assume that $p(s) = p(t) = 0$.  
%
  The DAG $\calG$ has a source node $\sigma \in \calV$ and a
  destination $\tau \in \calV$. Each node $\alpha \in \calV$ represent
  a VNF that has a processing demand $p(\alpha)$.
%
  Without loss of generality, we assume that $p(\sigma) = p(\tau) =
  0$.  Each edge $\eps \in \calE$ has a bandwidth demand $b(\eps)$.

@@

\begin{figure}[t]
\centering
\scalebox{0.9}{
\input{fig-embedding}
}
\caption{
  An example of an \scp instance and a solution.
  The left DAG represents the collection of alternative service chains as requested by the client.
  Each node of this graph represent a single service and the number written inside the node is the processing demand of this service.
  The value on the arcs are the minimum bandwidth required to transmit the output of one service to the next service.
  \\
  The right DAG represents a physical network. 
  The numbers on the nodes represent processing capacities and the number on the arcs represent the available bandwidth.
  \\
  The blue, dashed, arcs represent the chosen service chain and a physical route respectively, and the dotted, purple, arrows show a placement of this service chain into the physical route.
}
\label{fig:solution}
\end{figure}

@@

The \scp instance can be computed in polynomial time.
%
Note that any service chain is of the form $\sigma, \gamma_1, \dots, \gamma_n, \tau$ where $\gamma_i \in \{\alpha_i, \beta_i\}$. 
This implies that given a service chain, the sum of its cost and its total processing demand equals to $\sum_i a_i$ and thus (as forced by the processing capacity of $v$), $\set{a_1, \ldots, a_n} \in \textsc{Partition}$ if and only if there exists a solution whose cost is $\half \sum_i a_i$.
\end{proof}

@@

In this section we present an FPTAS for \scp in DAGs for general
costs, namely without the assumption that costs are polynomially
bounded.  Our algorithm is similar to the FPTAS for the
\textsc{Minimum Knapsack} problem~\cite{lawler1979fast}.

@@

\begin{theorem}
\label{thm:fptas}
There exists an FPTAS for \scp in DAGs.
\end{theorem}
\begin{proof}
Given $\eps>0$, run the dynamic programming algorithm from
Section~\ref{sub:DPL} $\abs{\calV} \cdot \abs{V}$ times, once for each
possible value of $c_{\max}$, and choose the best placement.
According to Lemma~\ref{lemma:guess} the best placement is
$(1+\eps)$-approximate.  The running time is polynomial in the input
size and in $1/\eps$.
\end{proof}

@@

We assume w.l.o.g that the physical network does not contain any leafs (beside, maybe, $s$ and $t$), otherwise, these nodes, for the purpose of our computation, can be safely removed from the network.

Our randomized algorithm consists of $k!$ iterations of the following
two phases: an orientation phase that applies a random orientation to
the physical network, and an execution of the FPTAS for DAGs given in
Theorem~\ref{thm:fptas}.  The algorithm finds a $(1+\eps)$-approximate
placement with probability $(1-1/e)$, and the running time of the
algorithm is $O((n+t(n))k!)$, where $t(n)$ is the time it takes to
compute a placement when the physical network is a DAG.
%
As usual, one may amplify the probability of success using repetition.

@@

\begin{theorem}
There exists an algorithm, that given an \scp instance with an optimal
solution that contains at most $k$ neighborly vertices, finds a
$(1+\eps)$-approximate placement with high probability, whose
running time is $O(k! \cdot \text{\rm poly}(n))$.
\end{theorem}
\begin{proof}
Consider an optimal placement with at most $k$ neighborly vertices,
and let $\pi'$ be the permutation it induces on the neighborly
vertices in the physical path.
%
If the random permutation $\pi$ orders the neighborly vertices
correctly, i.e., if $\pi(v) = \pi'(v)$, for every $v \in N$, then the
optimal placement is feasible with respect to $G_\pi$.  Hence, the
FPTAS will find a $(1+\eps)$-approximate placement in $G_\pi$.  This
placement is feasible, and therefore $(1+\eps)$-approximate placement,
with respect to $G$.
%
The permutation $\pi$ agrees with $\pi'$ with probability $1/k!$, and
the probability that $\pi$ agrees with $\pi'$ in $k!$ iterations is $1
- (1 - 1/k!)^{k!} \geq 1 - e^{-1}$.  
This means that the probability to fail in a single attempt is less than $e^{-1}$. If one makes $i$ attempts then the probability to fail in all those attempts is less than $e^{-i}$ so using $\ln n$ repetitions reduces the probability to fail to less than $\frac{1}{n}$ and increases the probability to succeed to at least $1 - \inv{n}$.
\end{proof}

@@

A biconnected component (block) of an undirected graph is a maximal biconnected subgraph.
Any connected graph decomposes into a tree of biconnected components called the block-cut tree of the graph. The blocks are attached to each other at shared vertices called cut vertices.
Computing such a tree can be done in linear time~\cite{HopcroftTarjan73}.
The orientation works as follow.  
Given an undirected graph $G = (V,E)$, we first compute its block-cut tree.
%
Then on the unique path from $s$ to $t$ in the block-cut tree we
internally orient each block arbitrarily to produce a DAG.  Note that
if $B_1, \ldots, B_\ell$ are the blocks of $G$ on the path from $s$ to
$t$ then the first and last vertices with respect to each block are
known and only the order of the internal neighborly vertices in each
block is needed to be determined, thus, we can find all relevant
orientations by repeating the process above $\Pi_{i = 1}^\ell
\abs{N_i}$ times, where $N_i$ is the set of internal neighborly
vertices in $B_i$.  See Figure~\ref{fig:or2} for an example.

@@

We say that two (physical) paths are \emph{internal vertex disjoint} if they do not have any internal (all vertices beside the first and last one) vertex in common.
Formally, given a physical network, a virtual specification, and a
parameter $r \in \naturals$, an \emph{$r$-robust} placement is a
collection of $r$ placements, $f_1,\ldots,f_r$, that correspond to $r$
internal vertex disjoint physical paths.  In the fault tolerant
version of \scp we are given $r$ costs functions $c_1,\ldots,c_r$, and
we are looking for an $r$-robust placement that minimize the total
cost:
\[
\sum_{i=1}^r c_i(f_i) = \sum_{i = 1}^r \sum_j c_i(\alpha_j,f_i(\alpha_j))
~.
\]
We refer to this problem as $r$-robust \scp.
%
The motivation for using different costs functions is that it may be
the case that the first placement is the primal placement that may be
replaced by one of the other placements in case of a malfunction along
the physical path that is used by $f_1$.  Hence, the cost function
$c_2$ may take into account the probability of using the placement
$f_2$.

@@

The algorithm uses dynamic programming.  We find the optimal 2-robust
placement for every combination of two physical vertices from the same
layer, and every combination of two virtual edges.
%
Formally, we define $\ft((\alpha_1,\zeta_1), (\alpha_2,\zeta_2), v_1,
v_2)$ to be the minimum cost 2-robust placement of two service chains
that start at $\sigma$ and end with the edges $(\alpha_1,\zeta_1)$ and
$(\alpha_2,\zeta_2)$, respectively, such that a virtual paths from
$\sigma$ to $\alpha_1$ and $\alpha_2$ are placed into internal vertex
disjoint physical paths that start at $s$ and end at $v_1$ and $v_2$,
respectively.
%
Recall that $\cost^i_{v_i}(\gamma_i,\beta_i)$ is the cost, with
respect to $c_i$, of a minimum cost path from $\gamma_i$ to $\beta_i$
whose total processing requirement can be handled by $v_i$.
%
Consider the following recursive definition.  First,
\[
\ft((\alpha_1,\zeta_1),(\alpha_2,\zeta_2),s,s)
= \cost^1_{s}(\sigma \leadsto \alpha_1) +
  \cost^2_{s}(\sigma \leadsto \alpha_2)
~.
\]
Otherwise, for $v_1 \neq v_2$, such that $v_1, v_2 \in L_i$, for some
$i$, we have
\[
\ft((\alpha_1,\zeta_1),(\alpha_2,\zeta_2),v_1,v_2)
=
\min\set{
%
\begin{array}{l}
\ft((\gamma_1,\beta_1),(\gamma_2,\beta_2),u_1,u_2) \\ 
+~ \cost^1_{v_1}(\beta_1 \leadsto \alpha_1) 
+~ \cost^2_{v_2}(\beta_2 \leadsto \alpha_2)
\end{array}
}
\]
where we take the minimum over all the possible virtual functions
$\gamma_1, \beta_1, \gamma_2, \beta_2$ and physical nodes $u_1, u_2$,
such that:
\begin{itemize}
  \item $\gamma_i \preceq \alpha_i$,
  \item $\beta_i \preceq \alpha_i$ or $\beta_i = \zeta_i$,
  \item $(\gamma_i,\beta_i) \in \calE$,
  \item $u_1 \neq u_2$ or $u_1 = u_2 = s$,
  \item $(u_i,v_i) \in E$, and
  \item $b(\gamma_i,\beta_i) \leq b(u_i,v_i)$
\end{itemize}

>>>>

% FIX
\newcommand{\fix}[1]{{\color{purple}\textbf{#1}}}
% DEBUG
%\usepackage{lipsum}
%\usepackage{todonotes}
%\usepackage{lineno}
%\linenumbers

@@

In this paper we consider an SDN network that employs
NFV~\cite{SDN-NFV15}.  Such networks attracted a lot of attention in
the networking community, especially from a systems point of view
(see, e.g.,~\cite{GVPGKDA14,HVSBFTF15}).  Several papers also
considered the algorithmic aspects of such networks but mainly
presented heuristics (see, e.g,~\cite{soule2014merlin}).  For more
details see~\cite{ERS16} and references therein.

@@

\subsection{Model}
%
An instance of the \scplong (\scp) problem is composed of three
components, a physical network, a virtual specification, and placement
costs:
\begin{description}
\item[Physical network:]
%
  The physical network is a graph $G = (V,E)$.  Each node $v \in V$
  has a non-negative processing capacity $p(v)$, and each directed
  edge $e \in E$ has a non-negative bandwidth capacity $b(e)$.
%
  % Without loss of generality, we assume that $p(s) = p(t) = 0$.

@@

\item[Virtual specification:]
%
  The description of a request for a service chain consists of a
  physical source $s \in V$, a physical destination $t \in V$, and a
  directed acyclic graph (DAG) $\calG = (\calV,\calE)$.
%
  The DAG $\calG$ has a source node $\sigma \in \calV$ and a
  destination $\tau \in \calV$. Each node $\alpha \in \calV$ represent
  a VNF that has a processing demand $p(\alpha)$.
%
  Without loss of generality, we assume that $p(s) = p(t) = p(\sigma) = p(\tau) =
  0$.  Each edge $\eps \in \calE$ has a bandwidth demand $b(\eps)$.

@@

\begin{figure}[t]
\centering
\scalebox{0.9}{
\input{fig-embedding}
}
\caption{\fix{
  An example of an \scp instance and a solution.
  The left DAG represents the collection of alternative service chains as requested by the client.
  Each node of this graph represent a single service and the number written inside the node is the processing demand of this service.
  The value on the arcs are the minimum bandwidth required to transmit the output of one service to the next service.
  \\
  The right DAG represents a physical network. 
  The numbers on the nodes represent processing capacities and the number on the arcs represent the available bandwidth.
  \\
  The blue, dashed, arcs represent the chosen service chain and a physical route respectively, and the dotted, purple, arrows show a placement of this service chain into the physical route.
}}
\label{fig:solution}
\end{figure}

@@

The \scp instance can be computed in polynomial time.
%
\fix{
Note that any service chain is of the form $\sigma, \gamma_1, \dots, \gamma_n, \tau$ where $\gamma_i \in \{\alpha_i, \beta_i\}$. 
This implies that given a service chain, the sum of its cost and its total processing demand equals to $\sum_i a_i$ and thus (as forced by the processing capacity of $v$), $\set{a_1, \ldots, a_n} \in \textsc{Partition}$ if and only if there exists a solution whose cost is $\half \sum_i a_i$.
}
\end{proof}

@@

In this section we present an FPTAS for \scp in DAGs for general
costs, namely without the assumption that costs are polynomially
bounded.  
\fix{
We use similar techniques as in the  FPTAS for the
\textsc{Minimum Knapsack} problem~\cite{islam2009approximation}.
}

@@

\begin{theorem}
\label{thm:fptas}
There exists an FPTAS for \scp in DAGs.
\end{theorem}
\begin{proof}
Given $\eps>0$, run the dynamic programming algorithm from
Section~\ref{sub:DPL} $\abs{\calV} \cdot \abs{V}$ times, once for each
possible value of $c_{\max}$, and choose the best placement.
\fix{Since we check all the possible values of $c_max$ and for each such value we compute an optimal solution with respect to $c'$, then} according to Lemma~\ref{lemma:guess} the best placement is $(1+\eps)$-approximate.  
The running time is polynomial in the input size and in $1/\eps$.
\end{proof}

@@

\fix{We assume w.l.o.g that the physical network does not contain any leafs (beside, maybe, $s$ and $t$), otherwise, these nodes, for the purpose of our computation, can be safely removed from the network.}
Our randomized algorithm consists of $k!$ iterations of the following
two phases: an orientation phase that applies a random orientation to
the physical network, and an execution of the FPTAS for DAGs given in
Theorem~\ref{thm:fptas}.  The algorithm finds a $(1+\eps)$-approximate
placement with probability $(1-1/e)$, and the running time of the
algorithm is $O((n+t(n))k!)$, where $t(n)$ is the time it takes to
compute a placement when the physical network is a DAG.
%
As usual, one may amplify the probability of success using repetition.

@@

\begin{theorem}
There exists an algorithm, that given an \scp instance with an optimal
solution that contains at most $k$ neighborly vertices, finds a
$(1+\eps)$-approximate placement with high probability, whose
running time is $O(k! \cdot \text{\rm poly}(n))$.
\end{theorem}
\begin{proof}
Consider an optimal placement with at most $k$ neighborly vertices,
and let $\pi'$ be the permutation it induces on the neighborly
vertices in the physical path.
%
If the random permutation $\pi$ orders the neighborly vertices
correctly, i.e., if $\pi(v) = \pi'(v)$, for every $v \in N$, then the
optimal placement is feasible with respect to $G_\pi$.  Hence, the
FPTAS will find a $(1+\eps)$-approximate placement in $G_\pi$.  This
placement is feasible, and therefore $(1+\eps)$-approximate placement,
with respect to $G$.
%
The permutation $\pi$ agrees with $\pi'$ with probability $1/k!$, and
the probability that $\pi$ agrees with $\pi'$ in $k!$ iterations is $1
- (1 - 1/k!)^{k!} \geq 1 - e^{-1}$.  
This means that the probability to fail in a single attempt is less than $e^{-1}$. \fix{If one makes $i$ attempts then the probability to fail in all those attempts is less than $e^{-i}$ so using $\ln n$ repetitions reduces the probability to fail to less than $\frac{1}{n}$ and increases the probability to succeed to at least $1 - \inv{n}$.
}
\end{proof}

@@

\fix{
A biconnected component (block) of an undirected graph is a maximal biconnected subgraph.
Any connected graph decomposes into a tree of biconnected components called the block-cut tree of the graph. The blocks are attached to each other at shared vertices called cut vertices.
Computing such a tree can be done in linear time~\cite{HopcroftTarjan73}.
}
The orientation works as follow.  
Given an undirected graph $G = (V,E)$, we first compute its block-cut tree.
%
Then on the unique path from $s$ to $t$ in the block-cut tree we
internally orient each block arbitrarily to produce a DAG.  Note that
if $B_1, \ldots, B_\ell$ are the blocks of $G$ on the path from $s$ to
$t$ then the first and last vertices with respect to each block are
known and only the order of the internal neighborly vertices in each
block is needed to be determined, thus, we can find all relevant
orientations by repeating the process above $\Pi_{i = 1}^\ell
\abs{N_i}$ times, where $N_i$ is the set of internal neighborly
vertices in $B_i$.  See Figure~\ref{fig:or2} for an example.

@@

\fix{We say that two (physical) paths are \emph{internal vertex disjoint} if they do not have any internal (all vertices beside the first and last one) vertex in common.}
Formally, given a physical network, a virtual specification, and a
parameter $r \in \naturals$, an \emph{$r$-robust} placement is a
collection of $r$ placements, $f_1,\ldots,f_r$, that correspond to $r$
internal vertex disjoint physical paths.  In the fault tolerant
version of \scp we are given $r$ costs functions $c_1,\ldots,c_r$, and
we are looking for an $r$-robust placement that minimize the total
cost:
\[
\sum_{i=1}^r c_i(f_i) = \sum_{i = 1}^r \sum_j c_i(\alpha_j,f_i(\alpha_j))
~.
\]
We refer to this problem as $r$-robust \scp.
%
The motivation for using different costs functions is that it may be
the case that the first placement is the primal placement that may be
replaced by one of the other placements in case of a malfunction along
the physical path that is used by $f_1$.  Hence, the cost function
$c_2$ may take into account the probability of using the placement
$f_2$.

@@

The algorithm uses dynamic programming.  We find the optimal 2-robust
placement for every combination of two physical vertices from the same
layer, and every combination of two virtual edges.
%
Formally, we define $\ft((\alpha_1,\zeta_1), (\alpha_2,\zeta_2), v_1,
v_2)$ to be the minimum cost 2-robust placement of two service chains
that start at $\sigma$ and end with the edges $(\alpha_1,\zeta_1)$ and
$(\alpha_2,\zeta_2)$, respectively, such that a virtual paths from
$\sigma$ to $\alpha_1$ and $\alpha_2$ are placed into internal vertex
disjoint physical paths that start at $s$ and end at $v_1$ and $v_2$,
respectively.
%
Recall that $\cost^i_{v_i}(\gamma_i,\beta_i)$ is the cost, with
respect to $c_i$, of a minimum cost path from $\gamma_i$ to $\beta_i$
whose total processing requirement can be handled by $v_i$.
%
Consider the following recursive definition.  First,
\[
\ft((\alpha_1,\zeta_1),(\alpha_2,\zeta_2),s,s)
= \cost^1_{s}(\sigma \leadsto \alpha_1) +
  \cost^2_{s}(\sigma \leadsto \alpha_2)
~.
\]
Otherwise, for $v_1 \neq v_2$ or $v_1 = v_2 = t$, such that $v_1, v_2 \in L_i$, for some
$i$, we have
\[
\ft((\alpha_1,\zeta_1),(\alpha_2,\zeta_2),v_1,v_2)
=
\min\set{
%
\begin{array}{l}
\ft((\gamma_1,\beta_1),(\gamma_2,\beta_2),u_1,u_2) \\ 
+~ \cost^1_{v_1}(\beta_1 \leadsto \alpha_1) 
+~ \cost^2_{v_2}(\beta_2 \leadsto \alpha_2)
\end{array}
}
\]
where we take the minimum over all the possible virtual functions
$\gamma_1, \beta_1, \gamma_2, \beta_2$ and physical nodes $u_1, u_2$,
such that:
\begin{itemize}
  \item $\gamma_i \preceq \alpha_i$,
  \item $\beta_i \preceq \alpha_i$ or $\beta_i = \zeta_i$,
  \item $(\gamma_i,\beta_i) \in \calE$,
  \item $u_1 \neq u_2$ or $u_1 = u_2 = s$,
  \item $(u_i,v_i) \in E$, and
  \item $b(\gamma_i,\beta_i) \leq b(u_i,v_i)$
\end{itemize}
